Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance

### 2017-07-19 ms_pre_post.Rmd
Initial attempt through methods

**Complex traits such as reproductive behaviour are likely to be highly polygenic [@Evans2008].** **Hence, in our model,** all individuals have 10 physically unlinked diploid loci, and therefore 20 alleles, underlying each of three reproductive strategy traits (i.e., 30 diploid loci in total): tendency for polyandry ($P_{a}$, i.e., 'polyandry' alleles), pre-copulatory inbreeding strategy ($M_{a}$, i.e., 'mating' alleles), and post-copulatory inbreeding strategy ($F_{a}$, i.e., 'fertilisation' alleles). All alleles can take the value of any real number [i.e., a continuum-of-alleles model; @Kimura1965; @Lande1976; @Reeve2000; @Bocedi2014] and combine additively to determine genotypes ($P_{g}$, $M_{g}$, and $F_{g}$) and resulting phenotypes ($P_{p}$, $M_{p}$, and $F_{p}$) for tendency for polyandry, pre-copulatory inbreeding strategy, and post-copulatory inbreeding strategy, respectively. Each individual's genotypic values $P_{g}$, $M_{g}$, and $F_{g}$ equal the sum of its 20 alleles for each trait. Each individual's phenotypic values for pre-copulatory and post-copulatory inbreeding strategy equal their respective genotypic values ($M_{p} = M_{g}$ and $F_{p} = F_{g}$), where negative and positive values cause inbreeding avoidance and preference, respectively (see details of mating and fertilisation strategies below). **In contrast to inbreeding strategy, the polygenic nature of polyandry does not translate to continuous phenotypes affecting mate choice [e.g., @Shuker2007; @Evans2013].** **Polyandry is more realistically defined as a ''threshold trait'', influenced by continuous genetic variation but expressed at some discrete phenotypic value(s) [@Lynch1998; @Roff1996; @Roff1998; @Duthie]. One important consequence of this threshold genetic architecture is that threshold traits can often reappear despite sustained negative selection as a consequence of previously masked deleterious alleles being reapparing due to recombination [@Roff1996; @Roff1998]. We incorporate this key feature into our model by allowing** phenotypic values for tendency for polyandry *to* equal genotypic values ($P_{p} = P_{g}$) only if $P_{g} \geq 0$, while $P_{p} = 0$ if $P_{g} < 0$ because females cannot choose a negative number of mates. Polyandry is therefore modelled as a threshold trait [@Lynch1998; @Roff1996; @Roff1998; @Duthie] that is influenced by continuous genetic variation but only expressed when $P_{g} > 0$. A negative $P_{g}$ value therefore generates a female that is phenotypically monandrous, while a positive $P_{g}$ value generates a female that can express some degree of polyandry (see details below).

@@

Phenotypic values of the three reproductive strategy traits each have costs that are assumed to combine independently to increase the probability that a focal female will die before mating. **The ability to set costs independently and a priori in this way is critical because it permits the direct comparison of evolution of pre-copulatory versus post-copulatory inbreeding avoidance** (see Discussion for consideration of other costs). The probabilities of pre-mating mortality due to the costs of polyandry ($c_{P}$), pre-copulatory inbreeding strategy ($c_{M}$), and post-copulatory inbreeding strategy ($c_{F}$) are $P_{p} \times c_{P}$, $|M_{p}| \times c_{M}$, and $|F_{p}| \times c_{F}$, respectively. Here $|M_{p}|$ and $|F_{p}|$ are the absolute values of $M_{p}$ and $F_{p}$, respectively. Absolute values are used for applying costs to inbreeding avoidance strategies because both negative and positive $M_{p}$ and $F_{p}$ values could potentially arise and affect the degree of inbreeding, representing inbreeding avoidance and inbreeding preference, respectively. In contrast, $P_{p}$ cannot be negative. **Values of $c_{P}$, $c_{M}$, and $c_{F}$ can be interpreted mechanistically as the increased mortality risk caused by mate searching [e.g., @Koga1998], resisting male harrassment [i.e., sexual conflict; @Rowe1994], and energetic loss from development of post-copulatory inbreeding avoidance mechanisms [e.g., @Gasparini2011], respectively.** Overall, because generations are non-overlapping, a female's probability of total reproductive failure increases linearly with the phenotypic value of each trait.

@@

To avoid unrestricted population growth, we set separate carrying capacities for the total numbers of females ($K_{f}$) and males ($K_{m}$) in the focal population following immigration [@Guillaume2009; @Duthie]. Hence, if at the end of a generation the number of female or male **offspring** exceeds $K_{f}$ or $K_{m}$  respectively, then individuals are randomly removed until each sex is at its carrying capacity. Such removal can be interpreted as some combination of dispersal and mortality. The remaining females and males form the next generation of potentially breeding adults.

@@

To address how asymmetric costs associated with pre-copulatory and post-copulatory inbreeding avoidance and polyandry affect the long-term persistence of reproductive strategies, we quantify the change in $M_{a}$ and $F_{a}$ over generations in simulations where pre-copulatory inbreeding strategy was cost free ($c_{M}=0$) but post-copulatory inbreeding strategy was moderately costly ($c_{F}=0.02$), and vice versa. We compare evolutionary trajectories with those of a costly strategy in the absence of evolution of an alternative strategy (e.g., evolution of pre-copulatory inbreeding strategy when post-copulatory inbreeding strategy phenotype is fixed at zero, $F_{p}=0$). Previous modelling using similar genetic architecture suggests that a cost of $0.02$ imposes strong but not overwhelming direct negative selection on polyandry [@Duthie]. This value is therefore appropriate to illustrate the different evolutionary consequences that could result from asymmetrical costs.

>>>>

**Complex traits such as reproductive behaviour are likely to be highly polygenic [@Evans2008].** **Hence, in our model,** all individuals have 10 physically unlinked diploid loci, and therefore 20 alleles, underlying each of three reproductive strategy traits (i.e., 30 diploid loci in total): tendency for polyandry ($P_{a}$, i.e., 'polyandry' alleles), pre-copulatory inbreeding strategy ($M_{a}$, i.e., 'mating' alleles), and post-copulatory inbreeding strategy ($F_{a}$, i.e., 'fertilisation' alleles). All alleles can take the value of any real number [i.e., a continuum-of-alleles model; @Kimura1965; @Lande1976; @Reeve2000; @Bocedi2014] and combine additively to determine genotypes ($P_{g}$, $M_{g}$, and $F_{g}$) and resulting phenotypes ($P_{p}$, $M_{p}$, and $F_{p}$) for tendency for polyandry, pre-copulatory inbreeding strategy, and post-copulatory inbreeding strategy, respectively. Each individual's genotypic values $P_{g}$, $M_{g}$, and $F_{g}$ equal the sum of its 20 alleles for each trait. Each individual's phenotypic values for pre-copulatory and post-copulatory inbreeding strategy equal their respective genotypic values ($M_{p} = M_{g}$ and $F_{p} = F_{g}$), where negative and positive values cause inbreeding avoidance and preference, respectively (see details of mating and fertilisation strategies below). **In contrast to inbreeding strategies, the polygenic nature of polyandry cannot translate to continuous phenotypes affecting mate choice [e.g., @Shuker2007; @Evans2013].** **Polyandry is more realistically defined as a "threshold trait", influenced by continuous genetic variation but expressed at some discrete phenotypic value(s) [@Lynch1998; @Roff1996; @Roff1998; @Duthie]. We incorporate this genetic architecture into our model by allowing** phenotypic values for tendency for polyandry **to** equal genotypic values ($P_{p} = P_{g}$) only if $P_{g} \geq 0$, while $P_{p} = 0$ if $P_{g} < 0$ because females cannot choose a negative number of mates. Polyandry is therefore influenced by continuous genetic variation but only expressed when $P_{g} > 0$. A negative $P_{g}$ value therefore generates a female that is phenotypically monandrous, while a positive $P_{g}$ value generates a female that can express some degree of polyandry (see details below). **One important consequence of this genetic architecture is that threshold traits often reappear in a population despite sustained negative selection, as caused by recombination of previously masked deleterious alleles [@Roff1996; @Roff1998].**

@@

Phenotypic values of the three reproductive strategy traits each have **pre-determined** costs that are assumed to combine independently to increase the probability that a focal female will die before mating. **The ability to set costs independently and a priori in this way is critical because it permits the direct comparison of evolution of pre-copulatory versus post-copulatory inbreeding avoidance** (see Discussion for consideration **of more mechanistic costs**). The probabilities of pre-mating mortality due to the costs of polyandry ($c_{P}$), pre-copulatory inbreeding strategy ($c_{M}$), and post-copulatory inbreeding strategy ($c_{F}$) are $P_{p} \times c_{P}$, $|M_{p}| \times c_{M}$, and $|F_{p}| \times c_{F}$, respectively. Here $|M_{p}|$ and $|F_{p}|$ are the absolute values of $M_{p}$ and $F_{p}$, respectively. Absolute values are used for applying costs to inbreeding avoidance strategies because both negative and positive $M_{p}$ and $F_{p}$ values could potentially arise and affect the degree of inbreeding, representing inbreeding avoidance and inbreeding preference, respectively. In contrast, $P_{p}$ cannot be negative. **Values of $c_{P}$, $c_{M}$, and $c_{F}$ can be interpreted mechanistically as the increased mortality risk caused by mate searching [e.g., @Koga1998], resisting male harrassment [i.e., sexual conflict; @Rowe1994], and energetic loss from development of post-copulatory inbreeding avoidance mechanisms [e.g., @Gasparini2011], respectively.** Overall, because generations are non-overlapping, a female's probability of total reproductive failure increases linearly with the phenotypic value of each trait.

@@

To avoid unrestricted population growth, we set separate carrying capacities for the total numbers of females ($K_{f}$) and males ($K_{m}$) in the focal population following immigration [@Guillaume2009; @Duthie]. Hence, if at the end of a generation the number of females or males exceeds $K_{f}$ or $K_{m}$  respectively, then individuals are randomly removed until each sex is at its carrying capacity. Such removal can be interpreted as some combination of dispersal and mortality. The remaining females and males form the next generation of potentially breeding adults.

@@

To address how asymmetric costs associated with pre-copulatory and post-copulatory inbreeding avoidance and polyandry affect the long-term persistence of reproductive strategies, we quantify the change in $M_{a}$ and $F_{a}$ over generations in simulations where pre-copulatory inbreeding strategy was cost free ($c_{M}=0$) but post-copulatory inbreeding strategy was moderately costly ($c_{F}=0.02$), and vice versa. We compare evolutionary trajectories with those of a costly strategy in the absence of evolution of an alternative strategy (e.g., evolution of pre-copulatory inbreeding strategy when post-copulatory inbreeding strategy phenotype is fixed at zero, $F_{p}=0$). Previous modelling using similar genetic architecture suggests that a cost of $0.02$ imposes strong but not overwhelming direct negative selection on polyandry [@Duthie]. This value is therefore appropriate to illustrate the different evolutionary consequences that could result from asymmetrical costs **(but see Supporting Information for other asymmetric cost combinations)**.
